---
title: "R Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 200
---

Estimating parameters of a population model using approximate Bayesian computation

Loading all necessary packages:

```{r genpack, message=FALSE, error=FALSE, warning=FALSE, echo=FALSE}
# install and load libraries if necessary:
genr_packs <- c("ggplot2", "gridExtra", "MASS", "dplyr", "coda", "grid")
for (package in genr_packs) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package, dependencies = TRUE)
  }
  else
    library(package, character.only = TRUE, quietly = TRUE)
}
```

```{r}
coral_data <- read.csv("data_project.csv")


summary_stats_table <- function(data) {
  mean_coral <- mean(data$coral)
  sd_coral <- sd(data$coral)
  mean_starfish <- mean(data$starfish)
  sd_starfish <- sd(data$starfish)
  
  summary_table <- data.frame(
    Species = c("Coral Reef", "Crown of Thorns Starfish"),
    Mean = c(mean_coral, mean_starfish),
    Std.Deviation = c(sd_coral, sd_starfish)
  )
  
  return(summary_table)
}

stats <- summary_stats_table(coral_data)
stats
```

```{r}
summary(coral_data)
```

```{r, fig.width=10}
ggplot(coral_data, aes(x = times)) +
  geom_line(aes(y=coral, color="Coral Reef"), linetype="solid", lwd=1) + 
  geom_point(aes(y=coral), size = 3, shape=19,  color="blue3") +
  geom_line(aes(y=starfish, color="Crown of Thorns Starfish"), linetype="dashed", lwd=1) + 
  geom_point(aes(y = starfish), size = 3, shape=18, color="red") +
  labs(title = "Coral and Starfish population over time (observed)", x = "Timeline (years)", y = "Species population") +
  scale_color_manual(values = c("Coral Reef" = "blue3", "Crown of Thorns Starfish" = "red")) +
  scale_x_continuous(breaks = coral_data$times) +
  theme()
```

Given, let $$C(t) \ \text{as the number of coral alive at time} \ t, and \\ S(t) \ \text{as the number of crown of thorn starfish alive at time} \ t.$$

The given transition probabilities represent the changes in the system in a small time interval $t+\Delta$ are:

$$
\begin{align*}
P(C = c + 1, \ S = s) & = \mu c \Delta_t + o(\Delta_t) \quad \text{(Birth event for coral)} \\
P(C = c - 1, \ S = s + 1) & = \delta s c \Delta_t + o(\Delta_t) \quad \text{(Predator event of starfish)} \\
P(C = c, \ S = s - 1) & = \nu s \Delta_t + o(\Delta_t) \quad \text{(Death event for starfish)}
\end{align*}
$$

$$\frac{dC}{dt} = \mu C − \delta CS$$ $$\frac{dS}{dt} = \nu S$$

C represents the prey population (coral). S represents the predator population (starfish). μ is the birth rate of the prey. δ is the predation rate, affecting both the prey and predator populations. ν
is the death rate of the predators.

Rate at which event happens: $\mu c \Delta_t + \delta s c \Delta_t + \nu s \Delta_t$

Birth with probability: $\mu c \Delta_t / \mu c \Delta_t + \delta s c \Delta_t + \nu s \Delta_t$ Predatory event with probability:
$delta s c \Delta_t / \mu c \Delta_t + \delta s c \Delta_t + \nu s \Delta_t$ Death with probability: $\nu s \Delta_t / \mu c \Delta_t + \delta s c \Delta_t + \nu s \Delta_t$

The inference task for this problem is to estimate the unknown parameters for the model, θ = (μ, δ, ν) using the observed data. The team has told you that uniform priors over zero to one would be
suitable for all parameters.

Let the observed data be denoted by y.

By Bayesian Inference, the posterior distribution: $$p(\theta | y) \propto p(y | \theta) \cdot p(\theta)$$ Likelihood function: $$P(y | \theta) = \prod_{i=1}^{n} P(C_i, S_i | \theta)$$ Prior
distributions: $$P(\mu) \sim U(0, 1),\\P(\delta) \sim U(0, 1),\\P(\nu) \sim U(0, 1)$$

Given some choices of θ, you can simulate data from this model using the stochastic simulation algorithm (SSA), commonly referred to as the Gillespie Algorithm.

```         
sim_Gillespie <- function(mu, delta, nu, C0 = 34, S0= 16, times = seq(from = 0,to = 20, by = 2)){
  x = c(C0, S0)
  states = matrix(0, nrow = length(times), length(x))
  states[1,] <- x

  for(i in 2:length(times)){
    x <- smfsb::stepLVc(x0 = x, t0 = times[i-1], deltat = times[i] - times[i-1], th = c(mu, delta, nu))
    states[i, ] <- x
  }
  return(data.frame(times=times, coral = states[,1], starfish = states[,2]))
}
```

```{r}
source("sim_Gillespie.R")
```

Exploring data with a single simulation:

```{r}
set.seed(1)
mu <- 0.6  # Birth rate for coral
delta <- 0.01  # Predator rate for starfish
nu <- 0.4  # Death rate for starfish

obsY <- sim_Gillespie(mu, delta, nu)
obsY
```

```{r}
par(mfrow=c(1,2))
plot(obsY$times, obsY$coral, type="o")
plot(obsY$times, obsY$starfish, type="o")
```

```{r, fig.width=11}
ggplot(obsY, aes(x = times)) +
  geom_line(aes(y=coral, color="Coral Reef (simulated)"), linetype="solid", lwd=1) + 
  geom_point(aes(y=coral), size = 3, shape=19,  color="blue3") +
  geom_line(aes(y=starfish, color="Crown of Thorns Starfish (simulated)"), linetype="dashed", lwd=1) + 
  geom_point(aes(y = starfish), size = 3, shape=18, color="red") +
  labs(title = "Coral and Starfish population (simulated)", x = "Timeline (years)", y = "Simulated species population") +
  scale_color_manual(values = c("Coral Reef (simulated)" = "blue3", "Crown of Thorns Starfish (simulated)" = "red")) +
  scale_x_continuous(breaks = obsY$times) +
  theme()
```

ABC Rejection

```{r, warning=False, fig.height=5, fig.width=15}
set.seed(1)
# Number of simulations
M <- 10000

# parameters of the model:
mu <- runif(M, 0, 1)
delta <- runif(M, 0, 1)
nu <- runif(M, 0, 1)

# parameter vector
theta <- matrix(c(mu, delta, nu), ncol = 3)
ret_index <- vector(length=M)
all_distances <- numeric()

for (i in 1:M) {
  sg_sim <- sim_Gillespie(theta[i,1], theta[i,2], theta[i,3])
  distances <- sqrt(sum((coral_data - sg_sim)^2))
  all_distances <- c(all_distances, distances)
}
eps <- quantile(all_distances, 0.05)
keeps <- all_distances < eps

prior_dist <- cbind(mu, delta, nu)
prior_df <- data.frame(mu = prior_dist[,1], delta = prior_dist[,2], nu = prior_dist[,3])

posterior_dist <- prior_dist[keeps,]
posteriors_df <- data.frame(mu = posterior_dist[,1], delta = posterior_dist[,2], nu = posterior_dist[,3])

plot_mu <- ggplot(prior_df,aes(x=mu)) +
  geom_density(color="red") +
  geom_density(data=posteriors_df,color="blue") + #geom_density(data=post_sdf,color="black") +
  geom_point(aes(x=0.75, y=0)) +
  xlab(expression(mu)) + ylab("density")

plot_delta <- ggplot(prior_df,aes(x=delta)) +
  geom_density(color="red") +
  geom_density(data=posteriors_df,color="blue") + #geom_density(data=post_sdf,color="black") +
  geom_point(aes(x=0.05, y=0)) +
  xlab(expression(delta)) + ylab("density")

plot_nu <- ggplot(prior_df,aes(x=nu)) +
  geom_density(color="red") +
  geom_density(data=posteriors_df,color="blue") + #geom_density(data=post_sdf,color="black") +
  geom_point(aes(x=0.03, y=0)) +
  xlab(expression(nu)) + ylab("density")

grid.arrange(plot_mu, plot_delta, plot_nu, ncol=3, nrow=1)
```

```{r}
set.seed(1)
# Number of simulations
M <- 10000

# parameters of the model:
mu <- runif(M, 0, 1)
delta <- runif(M, 0, 1)
nu <- runif(M, 0, 1)

# parameter vector
theta <- matrix(c(mu, delta, nu), ncol = 3)

theta_samples <- list()
# Simulating theta samples:
for (i in 1:M) {
  theta_samples[[i]] <- sim_Gillespie(theta[i,1], theta[i,2], theta[i,3]) #full data / no data summarisation 
}
#theta_samples_df <- do.call(rbind, theta_samples)

list_length <- length(theta_samples)
distances_real <- vector(length = M)
#distances_sample <- vector(length = M)

# Euclidean distances
for (i in 1:min(M, list_length)) {
  distances_real[i] <- sqrt(sum((coral_data - theta_samples[[i]])^2)) #coral_data
  #distances_sample[i] <- sum(obsY - theta_samples[[i]])^2 #simulated data
}

# Selecting top 5% of simulations
eps_r <- quantile(distances_real, 0.05, na.rm=TRUE)
keep_indR <- distances_real < eps_r

#eps_s <- quantile(distances_sample, 0.05, na.rm=TRUE)
#keep_indS <- distances_sample < eps_s

#parameters drawn from Prior
priors <- cbind(mu, delta, nu)
priors_df <- data.frame(mu = priors[,1], delta = priors[,2], nu = priors[,3])

#parameters drawn from Posterior
post_r <- priors[keep_indR,]
post_rdf <- data.frame(mu = post_r[,1], delta = post_r[,2], nu = post_r[,3])

#post_s <- priors[keep_indS,]
#post_sdf <- data.frame(mu = post_s[,1], delta = post_s[,2], nu = post_s[,3])
```


```{r, warning=False, fig.height=5, fig.width=15}
#Comparing prior and posterior:
plot_mu <- ggplot() +
  geom_density(data=priors_df, aes(x=mu), color="blue", linetype="dashed") +
  geom_density(data=post_rdf, aes(x=mu), color="red") + #geom_density(data=post_sdf,color="black") +
  geom_point(aes(x=0.75, y=0)) +
  xlab(expression(mu)) + ylab("density") +
  ggtitle("Posterior VS Prior for mu (coral birth rate)")

plot_delta <- ggplot() +
  geom_density(data=priors_df, aes(x=delta), color="blue", linetype="dashed") +
  geom_density(data=post_rdf, aes(x=delta), color="red") + #geom_density(data=post_sdf,color="black") +
  geom_point(aes(x=0.05, y=0)) +
  xlab(expression(delta)) +
  ggtitle("Posterior VS Prior for delta (predatory event)")

plot_nu <- ggplot() +
  geom_density(data=priors_df, aes(x=nu), color="blue", linetype="dashed") +
  geom_density(data=post_rdf, aes(x=nu), color="red") + #geom_density(data=post_sdf,color="black") +
  geom_point(aes(x=0.03, y=0)) +
  xlab(expression(nu)) +
  ggtitle("Posterior VS Prior for nu (starfish death rate)")

grid.arrange(plot_mu, plot_delta, plot_nu, ncol=3, nrow=1)
```

```{r}
mu_values <- post_rdf$mu
breaks <- seq(min(mu_values), max(mu_values), length.out = 20)
mu_bins <- cut(mu_values, breaks = breaks, include.lowest = TRUE)
mu_counts <- table(mu_bins)
most_common_MU <- names(mu_counts)[which.max(mu_counts)]
#mu_counts

del_values <- post_rdf$delta
breaks <- seq(min(del_values), max(del_values), length.out = 25)
del_bins <- cut(del_values, breaks = breaks, include.lowest = TRUE)
del_counts <- table(del_bins)
most_common_DEL <- names(del_counts)[which.max(del_counts)]
#del_counts

nu_values <- post_rdf$nu
breaks <- seq(min(nu_values), max(nu_values), length.out = 30)
nu_bins <- cut(nu_values, breaks = breaks, include.lowest = TRUE)
nu_counts <- table(nu_bins)
most_common_NU <- names(nu_counts)[which.max(nu_counts)]
#nu_counts

cat("Most common MU bin:", most_common_MU, "\n")
cat("Most common DEL bin:", most_common_DEL, "\n")
cat("Most common NU bin:", most_common_NU, "\n")
```

```{r}
summary(all_distances)
```

```{r}
summary(distances_real)
```

```{r, warning=FALSE, error=FALSE, fig.width=10, fig.height=5}

ggplot(data.frame(mu = post_rdf[,1], delta = post_rdf[,2], nu = post_rdf[,3], M = 1:length(post_rdf[,1])), aes(x = M)) +
  geom_line(aes(y = mu, color = "mu"), linetype = "solid") +
  geom_line(aes(y = delta, color = "delta"), linetype = "solid") +
  geom_line(aes(y = nu, color = "nu"), linetype = "solid") +
  ylab("") +
  xlab("Iteration") +
  scale_color_manual(name = "Parameters: ", values = c("mu" = "blue", "delta" = "orange", "nu" = "green3")) +
  theme(legend.position="right") +
  ggtitle("Parameter inference by ABC rejection")
```

```{r}
mcmc_abcr <- as.matrix(post_rdf)
mcmc_chains_abcr <- mcmc(mcmc_abcr)
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
#plot traceplots
par(mfrow=c(1,3))
traceplot(mcmc_chains_abcr[,1], main = paste("Trace Plot for mu"), col = "blue")
traceplot(mcmc_chains_abcr[,2], main = paste("Trace Plot for delta"), col = "orange")
traceplot(mcmc_chains_abcr[,3], main = paste("Trace Plot for nu"), col = "green3")
```

```{r, warning=FALSE, error=FALSE, fig.width=10, fig.height=5}
# Plot Lag
lag.plot(mcmc_chains_abcr[,1], main = "Lag Plot for mu", diag.col = "red")
lag.plot(mcmc_chains_abcr[,2], main = "Lag Plot for delta", diag.col = "red")
lag.plot(mcmc_chains_abcr[,3], main = "Lag Plot for nu", diag.col = "red")
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
# Plot ACF
par(mfrow=c(1,3))
acf(mcmc_chains_abcr[,1], main = paste("ACF for mu"))
acf(mcmc_chains_abcr[,2], main = paste("ACF for delta"))
acf(mcmc_chains_abcr[,3], main = paste("ACF for nu"))
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
# Plot Density
par(mfrow=c(1,3))
plot(density(mcmc_chains_abcr[,1]), lwd=2, col=4, main = paste("Density Plot for mu"))
plot(density(mcmc_chains_abcr[,2]), lwd=2, col=7, main = paste("Density Plot for delta"))
plot(density(mcmc_chains_abcr[,3]), lwd=2, col=3, main = paste("Density Plot for nu"))
```

```{r, fig.height=5, fig.width=10}
df_list <- list()

# Generate 100 data frames with random data
for (i in 1:100) {
  g <- sim_Gillespie(post_rdf[i,1], post_rdf[i,2], post_rdf[i,3])
  df_list[[i]] <- g
}

combined_df <- do.call(rbind, df_list)

ggplot(combined_df, aes(x = times)) +
  geom_line(aes(y=coral, color="Coral Reef (simulated)"), linetype="solid", lwd=1, alpha=0.3) + 
  geom_point(aes(y=coral), size = 2, shape=19,  color="blue3") +
  geom_line(aes(y=starfish, color="Crown of Thorns Starfish (simulated)"), linetype="dashed", lwd=1, alpha=0.3) + 
  geom_point(aes(y = starfish), size = 2, shape=18, color="red") +
  labs(title = "Coral and Starfish population (simulated)", x = "Timeline (years)", y = "Simulated species population") +
  scale_color_manual(values = c("Coral Reef (simulated)" = "blue3", "Crown of Thorns Starfish (simulated)" = "red")) +
  scale_x_continuous(breaks = combined_df$times) +
  theme()
```



**MCMC ABC**

```{r}
#post_rdf

df_nona <- post_rdf %>%
  mutate(across(everything(), ~ifelse(is.na(.), 0, .)))

log_values <- log(df_nona[, c(1, 2, 3)] / (1 - df_nona[, c(1, 2, 3)]))
transp_theta <- as.data.frame(log_values)

covm_post <- cov(transp_theta)
covm_post
cov_list <- c(covm_post)
```

```{r}
theta_current <- c(0.65, 0.04, 0.03)
cov_list
cov_2rw <- matrix(cov_list, 3, 3)

theta_trans_ct <- c(log((theta_current[1])/(1-theta_current[1])), 
                      log(theta_current[2]/(1-theta_current[2])), 
                      log(theta_current[3]/(1-theta_current[3])))
theta_trans_ct

theta_trans_pr <- mvrnorm(1,theta_trans_ct, cov_2rw)
theta_trans_pr
```

```{r ABCMCMC, include=FALSE}
set.seed(1)

M <- 10000 # Number of MCMC iterations
tol <- 316 # ABC tolerance

theta_current <- c(0.65, 0.04, 0.03)
cov_2rw <- matrix(cov_list, 3, 3)

samps <- matrix(0, M, 3)

for (i in 1:M) { # propose on the transformed space
  #print(theta_current)
  theta_trans_ct <- c(log(theta_current[1]/(1 - theta_current[1])), 
                      log(theta_current[2]/(1 - theta_current[2])), 
                      log(theta_current[3]/(1 - theta_current[3])))
  
  theta_trans_pr <- mvrnorm(1, theta_trans_ct, cov_2rw)

  theta_prop <- c(1/(1 + exp(-theta_trans_pr[1])), 
                  1/(1 + exp(-theta_trans_pr[2])), 
                  1/(1 + exp(-theta_trans_pr[3])))

  # compute prior densities on transformed space
  log_prior_curr <- -theta_trans_ct[1] - 2*log(1 + exp(-theta_trans_ct[1])) - 
    theta_trans_ct[2] - 2*log(1 + exp(-theta_trans_ct[2])) - 
    theta_trans_ct[3] - 2*log(1 + exp(-theta_trans_ct[3]))

  log_prior_prop <- -theta_trans_pr[1] - 2*log(1 + exp(-theta_trans_pr[1])) - 
    theta_trans_pr[2] - 2*log(1 + exp(-theta_trans_pr[2])) - 
    theta_trans_pr[3] - 2*log(1 + exp(-theta_trans_pr[3]))
  
  
  u <- runif(1)
  if (u > exp(log_prior_prop - log_prior_curr)) { # reject early
    samps[i,] <- theta_current
    next # skipping to next iteration
  }
  
  # Simulate data from the model
  x <- sim_Gillespie(theta_prop[1], theta_prop[2], theta_prop[3])
  distances <- sqrt(sum((coral_data - x)^2))
  #distances <- ifelse(is.na(distances), 0, distances)
  if (distances <= tol) {
    theta_current <- theta_prop
  }
  samps[i,] = theta_current
  print(i)
}
```

```{r}
dim(samps)
```

```{r}
# Check number samples
coda::effectiveSize(samps)
```

```{r}
# Compare rejection and MCMC ABC
postr_abc_df <- data.frame(mu = samps[seq(1,M,10),1], 
                           delta = samps[seq(1,M,10),2], 
                           nu = samps[seq(1,M,10),3])

postr_abc_df
```

```{r warning=False, fig.height=5, fig.width=15}
plot_mc_p <- ggplot() + 
  geom_density(data=postr_abc_df, aes(x=mu), color="red") + 
  geom_density(data=post_rdf, aes(x=mu), color="blue", linetype="dashed") +
  geom_density(data=priors_df, aes(x=mu), color="orange3", linetype="dashed") +
  theme_minimal() +
  geom_point(aes(x=0.75, y=0)) +
  xlab(expression(mu)) + ylab("density") +
  ggtitle("Posterior distributions for mu")

plot_mc_mu <- ggplot() +
  geom_density(data=postr_abc_df, aes(x=delta), color="red") +
  geom_density(data=post_rdf, aes(x=delta), color="blue", linetype="dotdash") +
  geom_density(data=priors_df, aes(x=delta), color="orange3", linetype="dashed") +
  theme_minimal() +
  geom_point(aes(x=0.05, y=0)) +
  xlab(expression(delta)) +
  ggtitle("Posterior distributions for delta")

plot_mc_phi <- ggplot() +
  geom_density(data=postr_abc_df, aes(x=nu), color="red") +
  geom_density(data=post_rdf, aes(x=nu), color="blue", linetype="dashed") +
  geom_density(data=priors_df, aes(x=nu), color="orange3", linetype="dashed") +
  theme_minimal() +
  geom_point(aes(x=0.03, y=0)) +
  xlab(expression(nu))+
  ggtitle("Posterior distributions for nu")

grid.arrange(plot_mc_p, plot_mc_mu, plot_mc_phi, ncol=3, nrow=1, top = textGrob("ABC rejection VS ABC MCMC", gp=gpar(fontsize=16,font=2)))
```

```{r, warning=FALSE, error=FALSE, fig.width=10, fig.height=5}
# Plot MCMC samples
ggplot(data.frame(mu = samps[,1], delta = samps[,2], nu = samps[,3], M = 1:length(samps[,1])), aes(x = M)) +
  geom_line(aes(y = mu, color = "mu"), linetype = "solid") +
  geom_line(aes(y = delta, color = "delta"), linetype = "solid") +
  geom_line(aes(y = nu, color = "nu"), linetype = "solid") +
  ylab("") +
  xlab("Iteration") +
  scale_color_manual(name = "Parameters: ", values = c("mu" = "blue", "delta" = "orange", "nu" = "green3")) +
  theme(legend.position="bottom")
```

```{r}
mcmc_abc <- as.matrix(samps)
mcmc_chains_abc <- mcmc(mcmc_abc)
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
par(mfrow=c(1,3))
traceplot(mcmc_chains_abc[,1], main = paste("Trace Plot for mu"), col = "blue")
traceplot(mcmc_chains_abc[,2], main = paste("Trace Plot for delta"), col = "orange")
traceplot(mcmc_chains_abc[,3], main = paste("Trace Plot for nu"), col = "green3")
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
# Plot Lag
lag.plot(mcmc_chains_abc[,1], main = "Lag Plot for mu", diag.col = "red", )
lag.plot(mcmc_chains_abc[,2], main = "Lag Plot for delta", diag.col = "red")
lag.plot(mcmc_chains_abc[,3], main = "Lag Plot for nu", diag.col = "red")
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
# Plot ACF
par(mfrow=c(1,3))
acf(mcmc_chains_abc[,1], main = paste("ACF for mu"))
acf(mcmc_chains_abc[,2], main = paste("ACF for delta"))
acf(mcmc_chains_abc[,3], main = paste("ACF for nu"))
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
# Plot Density
par(mfrow=c(1,3))
plot(density(mcmc_chains_abc[,1]), lwd=2, col=4, main = paste("Density Plot for mu"))
plot(density(mcmc_chains_abc[,2]), lwd=2, col=7, main = paste("Density Plot for delta"))
plot(density(mcmc_chains_abc[,3]), lwd=2, col=3, main = paste("Density Plot for nu"))
```


**SMC ABC MCMC**

```{r}
# Function to compute the prior density on transformed space
log_prior <- function(theta_trans) {
  return(-theta_trans[1] - 2 * log(1 + exp(-theta_trans[1])) - 
           theta_trans[2] - 2 * log(1 + exp(-theta_trans[2])) - 
           theta_trans[3] - 2 * log(1 + exp(-theta_trans[3])))
}

# Adaptive SMC Function
adaptive_smc_abc <- function(M, tau, tol_init, theta_init, cov_2rw, obs_data) {
  samps <- matrix(0, M, 3)
  weights <- rep(1/M, M)
  epsilons <- rep(tol_init, tau)
  indkeep <- sample(1:M, M, replace = TRUE, prob = weights)
  
  for (t in 1:tau) {
    theta_trans_ct <- log(theta_init/(1 - theta_init))
    theta_trans_pr <- mvrnorm(M, theta_trans_ct, cov_2rw)
    theta_prop <- 1/(1 + exp(-theta_trans_pr))
    
    # Compute prior densities on transformed space
    log_prior_curr <- log_prior(theta_trans_ct)
    log_prior_prop <- log_prior(theta_trans_pr)
    
    # Simulate data from the model
    distances <- apply(theta_prop, 1, function(theta) {
      sim_data <- sim_Gillespie(theta[1], theta[2], theta[3])
      sqrt(sum((obs_data - sim_data)^2))
    })
    
    epsilons[t] <- median(distances)
    
    ESS <- 1 / sum(weights^2) #adaptation step
    print(ESS)
    if (ESS < 316) { # Resample
      indkeep <- sample(1:M, M, replace = TRUE, prob = weights)
      weights <- rep(1/M, M)
    }
    
    weights <- exp(-distances^2 / (2 * epsilons[t]^2))
    weights <- weights / sum(weights)
    samps <- theta_prop[indkeep,] #store samples
    
    if (t < tau) { # Update tolerance for the next iteration
      quantile_75 <- quantile(distances, 0.05)
      tol_scaling_factor <- 7
      epsilons[t + 1] <- min(quantile_75 * tol_scaling_factor, tol_init)
    }
  }
  return(list(samples = samps, weights = weights, epsilons = epsilons))
}
```


```{r}
set.seed(1)
M <- 1000
tau <- 10
tol_init <- 316
theta_init <- c(0.3, 0.8, 0.3)
cov_2rw <- matrix(cov_list, 3, 3)

abc_smc_result <- adaptive_smc_abc(M, tau, tol_init, theta_init, cov_2rw, coral_data)

samps_smc <- abc_smc_result$samples
#abc_smc_result$weights
#abc_smc_result$epsilons
```



```{r, fig.width=10}
# Plot MCMC samples
ggplot(data.frame(mu = samps_smc[,1], delta = samps_smc[,2], nu = samps_smc[,3], M = 1:length(samps_smc[,1])), aes(x = M)) +
  geom_line(aes(y = mu, color = "mu"), linetype = "solid") +
  geom_line(aes(y = delta, color = "delta"), linetype = "solid") +
  geom_line(aes(y = nu, color = "nu"), linetype = "solid") +
  ylab("") +
  xlab("Iteration") +
  scale_color_manual(name = "Parameters: ", values = c("mu" = "blue", "delta" = "orange", "nu" = "green3")) +
  theme(legend.position="bottom")
```

```{r}
mcmc_abc_smc <- as.matrix(samps_smc)
mcmc_chains_abcsmc <- mcmc(mcmc_abc_smc)
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
par(mfrow=c(1,3))
traceplot(mcmc_chains_abcsmc[,1], main = paste("Trace Plot for mu"), col = "blue")
traceplot(mcmc_chains_abcsmc[,2], main = paste("Trace Plot for delta"), col = "orange")
traceplot(mcmc_chains_abcsmc[,3], main = paste("Trace Plot for nu"), col = "green3")
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
# Plot ACF
par(mfrow=c(1,3))
acf(mcmc_chains_abcsmc[,1], main = paste("ACF for mu"))
acf(mcmc_chains_abcsmc[,2], main = paste("ACF for delta"))
acf(mcmc_chains_abcsmc[,3], main = paste("ACF for nu"))
```

```{r, warning=FALSE, error=FALSE, fig.width=15, fig.height=5}
# Plot Density
par(mfrow=c(1,3))
plot(density(mcmc_chains_abcsmc[,1]), lwd=2, col=4, main = paste("Density Plot for mu"))
plot(density(mcmc_chains_abcsmc[,2]), lwd=2, col=7, main = paste("Density Plot for delta"))
plot(density(mcmc_chains_abcsmc[,3]), lwd=2, col=3, main = paste("Density Plot for nu"))
```

```{r}


```


